---
title: "True Fixed Effect Stochastic Frontier Models Estimation via Integrated Likelihood: ilsfa package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{il_sfa_Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ilsfa)
```

The package uses `R` and `Rcpp` code to manage the estimation of True Fixed Effect Stochastic Frontier 
Model (TFE SFM) via Integrated Likelihood (IL). In particular, 
the code includes two principal functions that are fully described in the package help documentation, and 
a bunch of background functions (most of them written in `Rcpp`) used to make the code computationally 
efficient. The present vignette focuses on the usage of

- `il_sfa` function for the model estimation and
- `DGP` function for the data generating process adopted in the simulation studies (see Bellio and Grassetti, 2023).

## Estimating TFE SFM via IL

The il_sfa function allows estimating the frontier model defined in Belotti and Ilardi (2018) and 
Bellio and Grassetti (2023), by using a feasible, generalizable and computationally efficient approach 
based on the integrated likelihood. In particular, the integrated likelihood approach works 
integrating-out the panel-specific intercepts from the likelihood function, solving the
incidental parameters issue without loss in estimation efficiency.

For a production frontier model,  the specification of the true fixed effects model is given by

\begin{equation}
\label{spec}
y_{it} =  \alpha_i + x_{it}^\top\beta  - u_{it} + \nu_{it} \, ,
\end{equation}
where $i$ is the index for panel ($i=1,\ldots,n$) and $t$ is the index for time 
($t=1,\ldots,T)$, with the panel size $T$ assumed to be the same across panels without loss of generality. 
Here $y_{iy}$ is the  output and $x_{it}$ a vector of exogenous inputs, both possibly
in log form, 
$\beta$ is a vector of structural coefficients and $\alpha_i$ the panel fixed effect; in the following, we 
will also employ $y_i$ and $X_i$ to denote the vector of outputs and the matrix of 
inputs for the $i$-th panel, respectively.
Finally, $u_{it}$ is the one-sided inefficiency term, and $v_{it}$ is the 
symmetric idiosyncratic error, independent of $u_{it}$. A simple change in the $u_{it}$ sign is required 
for a cost frontier model. 

The theory for composite group families \citep[Chap. 7]{pace1997principles} implies that a location parameter  might be eliminated simply by integrating it, with no loss of information. In other words, a 
general marginal log-likelihood for $\theta = \left\{\beta, \sigma\right\}$ can be defined as
\begin{equation}
\label{inte}
\ell_{\rm M}(\theta) = \sum_{i=1}^n \log \int_{-\infty}^\infty f(y_i;\alpha_i, \theta) d \alpha_ i \, ,
\end{equation}
where $f(y_i;\alpha_i, \theta)$ is the density function of
$y_i$.

Mimicking the Data Generating Process introduced in Belotti and Ilardi (2018) we can generate one single 
dataset with the setting used in Table 1, namely $N=100$, $T=5$, $\beta = 1$, $\gamma_0 = -1.5$,
$\gamma_1 = 1$, and $\sigma_v = 0.25$. A small number of replications (10) are considered here.

```{r }
N <- 100
T <- 5
para <- c(1,-1.5,1,0.25)
dat <- DGP(N, T, list(beta=para[1], sigmav=para[4], gamma=para[2:3]), model = "Exp",
           seed.val = 22, nsim = 10)
```

The estimation function needs a set of mandatory arguments, including all the data used in model specification. 

- a vector for the response variable,
- a matrix for the independent variables,
- a vector for the identifier of groups (cross-sectional units), and 
- a vector for the variable influencing the variance of the efficiency-specific error term.

These objects are here collected into a list for convenience, but this is not necessary.

```{r }
mydat <- list(y = dat$daty[1,], x = data.frame(dat$x), g = dat$g, z = dat$datz)
```

The estimation result is obtained here considering the following additional arguments.

- `distr` to specify the distribution of the efficiency-related error term. 
`Exp` stands for Exponential distribution.
- `het` to specify the presence of heteroscedasticity affecting the efficiency-related 
error term. `TRUE` means that the variance term $\sigma_{iu}$ is defined as $e^{\gamma_0 + \gamma_1z_it}$.
- `useHess` to manage the initialization of the maximization procedure. `TRUE` means 
that the likelihood maximization is initialized with the numerical Hessian; otherwise, 
the Hessian is initialized as a unit diagonal.

```{r }
mle <- il_sfa(X = mydat$x, y = mydat$y, distr = "Exp", het = TRUE, z = mydat$z,
              group = mydat$g, useHess = TRUE)
summary(mle)
```

The object created by `il_sfa` includes:

- `par` the parameters' point estimates,
- `invess` the inverse of the numerical hessian used to compute the standard errors of the estimated parameters,
- `se` the standard errors of the estimated parameters,
- `alphai` the firm-specific effects -- $\alpha_i$
- `uit` the estimated efficiency-related effects.

In some specific cases, the function fails, when considering the numerical hessian in 
the initialization of the maximization procedure. The estimation routine considers 
some checks on this issue, adopting the unit diagonal solution when the initialization 
fails. 

In the case of `HN` model estimation, the heteroscedastic efficiency-related errors 
are still not supported. The estimation routine again considers some preliminary checks 
and operates a feasible estimation sending a warning message. 

```{r }
mle <- il_sfa(X = mydat$x, y = mydat$y, distr = "HN", het = TRUE, z = mydat$z,
              group = mydat$g, useHess = TRUE)
```

## Developing a simulation study using ilsfa and DGP functions

The DGP function can be used to generate data for a simulation study. 
The following steps can be used to reproduce the results in Table 1 panel (a)
Belotti and Ilardi (2018). In particular, the simulation is the one considering 
$N=100$ and $T=5$. 

The example considers 10 replications only to show that the code is quite efficient from 
a computational point of view. Changing the `nsim` parameter, one can consider a different 
number of replications (the original results are obtained using `nsim = 1000`). 

```{r }
N <- 100
T <- 5
para <- c(1,-1.5,1,0.25)
nsimul <- 10
dat <- DGP(N, T, list(beta=para[1], sigmav=para[4], gamma=para[2:3]), model = "Exp",
           seed.val = 22, nsim = nsimul)
```

The estimation results can then be obtained on the simulated dataset saving the point estimates only.

```{r }
res <- matrix(0, nrow = nsimul, ncol=4)
for(i in 1:nrow(res)){
#  print(i)
  mydat <- list(y = dat$daty[i,], x = data.frame(dat$x), g = dat$g, z = dat$datz)
  mle <- il_sfa(X = mydat$x, y = mydat$y, distr = "Exp", het = TRUE, z = mydat$z,
                group = mydat$g, useHess = TRUE)
  res[i,] <- mle$par
  }
```

The obtained matrix can be used to compute the Mean Square Error (MSE) using a function included in the `R` source code.

```{r }
for(i in 1:4) print(MSE(res[,i], para[i]))
```

To check the goodness of our proposal, we also compared the estimation results with a 
benchmark identified with the results of Belotti and Illardi (2018). These results are 
obtained for the same generated datasets in Stata using `sftfe` function and adopting 
the `PDE` method. The results are collected in the package dataset called `res_N100_Stata.rda`.

```{r }
data(res_N100_Stata)

for(i in 1:4) print(MSE(res_N100_Stata[1:nsimul,i], para[i]))
```

The estimation results are quite similar, but the IL MSE are always a bit lower than the PDE ones.


## References

- Belotti, F., & Ilardi, G. (2018). Consistent inference in fixed-effects stochastic frontier models. *Journal of Econometrics*, **202**(2), 161–-177.
- Belotti, F., Daidone, S., Ilardi, G., Atella, V., 2013. Stochastic frontier analysis using Stata. The Stata Journal **13**, 719--758.
- Bellio, R., & Grassetti, L. (2023). Efficient estimation of true fixed-effects stochastic frontier models. *Working Paper*.
